Databrary: A digital data library for sharing research video

### 2015-01-20 jcdl-15/paper.tex
removing notes
wording

\begin{abstract}
%%dg being that the conference them is big data, should we be trying to be more explicit about that?
% big data is generally referring to "more rows" rather than larger files. I think the closest we get is the size of video files
% which is already noted below. Just wondering if it makes sense to be more explicit or not.
%%dg also preservation and discoverability - though maybe that's included in data management
%%rg video is big/dense in time and space. But, much of the size is masked without intensive analysis by humans or machines. Compare it with brain imaging data. A typical brain imaging study might include 3-5 runs of 200 brain volumes with 30 128x128 pixel x 8 bit greyscale slices per volume, with each volume collected about every 2-3 sec (.33-.5 Hz). Video is much "bigger" because it samples space and time densely.

\end{abstract}

@@

%%dg maybe break this part up into subsections: Content/Domain; Users; Design contraints; Approach (or something like that)

@@

%\subsection{Affordances and Challenges of Video}

%%dg perhaps better to frame below as the affordances and challenges of working with video data. 
%That is..."A key component to understanding the design decisions made in creating Databrary is the unique affordances and challenges of video and audio data"
%rog The text you removed explained how and why video is actually "data." Maybe that's not important for this audience. What are the questions we are trying to answer with this paper? I'd argue for the following: 1) Why video? 2) What challenges does video pose? 3) How does Databrary solve these challenges? 4) What challenges remain unmet? 5) What lessons does Databrary offer other data repositories?
%Affordances
We have intentionally chosen to focus our efforts on hosting human-consumable, self-documenting, timeseries data: video and audio recordings.
%rog I don't know that many people understand what the term affordances means.
These types of data have unique affordances but also present additional challenges \cite{Lanagan_Smeaton_2012}.
On one hand, recordings can be and largely are consumed and analyzed, at least initially, without sophisticated tools: once digitally decoded, trained human observers can glean directly much of the data's richness.
In particular, video data collected in one context can often be used by others for different purposes with minimal explanation, as the nature of the subjects and activity are readily apparent.
This naturally reduces the problems of documentation and re-use that plague most flat-file or dense datasets.
%Challenges
On the other hand, since automated analysis of videos is an active yet immature research area, the problems of search, discoverability, and meta-analysis over these media remain \cite{Albertson_2013}.
We attempt to address these issues by augmenting the primary recording data with additional structured and unstructured features.
%%dg probably need to make this section more cohesive, but the main gist is 1) We focus on video data, 2)video data has these inherent challenges. 3) At the same time there are already challenges involved in creating a research data repository, independant of the format of the data travels on. 4) This paper dicusses the challenges we have resolved and the challenges we have coming up.

@@

%%dg would a figure illustration be helpful here? do we already have one?
%%dg in general, how much room do we have for screenshots, if they help explain?

@@

% should we actually use the term "record"?  Alternatives: "grouping", "label", "annotation", ...
% as long as we're clear in defining it. the record section appears to be doing this.

@@

%dg "" - stuff like this to remind the reader why we're discussing these components in specific (rather than some other aspect of the system).
Researchers record activities with individuals either for the purpose of a later, closer review or analysis, or simply to document and verify measurements and observations made during the collection.
Thus, the recording itself is often the core item of research, representing a single data collection, experiment, trial, visit, or case depending on the nature of the research.
However, in many cases a single session can comprise multiple videos: from serial data streams due to recording equipment producing multiple files, and/or in parallel from multiple data capture sources operating at the same time.
These primary recordings may also be augmented with other forms of timeseries data such as motion capture, eye tracking, or physiological recordings, as well as with additional point-in-time measurements, observations, or surveys, all of which have specific and critical temporal relationships with the primary data.

@@

A session object captures these relationships to represent a single data acquisition, comprising not only arbitrary files but also their absolute or relative temporal positioning.
The metaphor we use here is a timeline, wherein each session defines a time range, with absolute timestamps representing the exact time of data collection, or relative positioning defined by a 0 starting point of the collection (or potentially any other meaningful wall-time units) with an optional date.
Files placed within a session can then be (optionally) positioned, much as in video or audio editing software.
Although some advanced recording equipment provides automatic timestamping or synchronization than can be used, most researchers collecting multiple parallel data streams use ``sync points'' such as an electronic trigger signal, flash, or tone to align recordings.
These sync points can then be used to manually place recordings on the timeline appropriately.

@@

%%dg maybe add some langauge to the beginning of this section that is explicit that the "record" is designed with the flexibility we talk about above in mind.
An object called a record serves a number of purposes around session and metadata organization. 
A record itself represents any number of variables or measurements---what would traditionally be a row of a database or flat file.
For example, a record may represent an individual participant within the dataset and contain their subject identifier, gender, birth date, and any other demographic, survey, or measurement data about that individual.
A record could also represent a particular experimental task, condition, location, or outcome, and contain variables with corresponding parameters or descriptions.
Records are thus labeled and divided by the type of entity they represent.
The variables themselves are textual, numeric, date, or potentially any other scalar data as in any traditional tabular data interface, with the records of a given type then constituting the tabular rows.

@@

%%dg other notes: should we give a bit more discussion of the infrastruction - Postgresql, ffmpeg, using HPC for transcoding, storage, frontend technologies etc. 
% i'm not sure. do other papers have this sort of "methods" section? -dylan
%%dg - other papers in this general area (see: Peer_2012, Griffin_etal_2014, and Witt_2012 from references) do comment on the technology used (if there actually is a product and not just a proposed framework) to varying degrees. I don't think we have to spend more than a paragraph on it, but it would be good to mention the stack in terms of taking, augmenting, and storing video files and metadata overall.

>>>>

\begin{abstract}
%%dg being that the conference them is big data, should we be trying to be more explicit about that?
% big data is generally referring to "more rows" rather than larger files. I think the closest we get is the size of video files
% which is already noted below. Just wondering if it makes sense to be more explicit or not.
%%dg also preservation and discoverability - though maybe that's included in data management
%%rg video is big/dense in time and space. But, much of the size is masked without intensive analysis by humans or machines. Compare it with brain imaging data. A typical brain imaging study might include 3-5 runs of 200 brain volumes with 30 128x128 pixel x 8 bit greyscale slices per volume, with each volume collected about every 2-3 sec (.33-.5 Hz). Video is much "bigger" because it samples space and time densely.
\end{abstract}

@@

@@

%\subsection{Benefits and Challenges of Video}

%rog The text you removed explained how and why video is actually "data." Maybe that's not important for this audience. What are the questions we are trying to answer with this paper? I'd argue for the following: 1) Why video? 2) What challenges does video pose? 3) How does Databrary solve these challenges? 4) What challenges remain unmet? 5) What lessons does Databrary offer other data repositories?
%Benefitss
We have intentionally chosen to focus our efforts on hosting human-consumable, self-documenting, timeseries data: video and audio recordings.
Though while these types of data are an accessible and potent source for research analysis, they also present additional challenges \cite{Lanagan_Smeaton_2012}.
On one hand, recordings can be and largely are consumed and analyzed, at least initially, without sophisticated tools: once digitally decoded, trained human observers can glean directly much of the data's richness.
In particular, video data collected in one context can often be used by others for different purposes with minimal explanation, as the nature of the subjects and activity are readily apparent.
This naturally reduces the problems of documentation and re-use that plague most flat-file or dense datasets.
%Challenges
On the other hand, since automated analysis of videos is an active yet immature research area, the problems of search, discoverability, and meta-analysis over these media remain \cite{Albertson_2013}.
We attempt to address these issues by augmenting the primary recording data with additional structured and unstructured features.
%%dg probably need to make this section more cohesive, but the main gist is 1) We focus on video data, 2)video data has these inherent challenges. 3) At the same time there are already challenges involved in creating a research data repository, independant of the format of the data travels on. 4) This paper dicusses the challenges we have resolved and the challenges we have coming up.

@@

%%dg would a figure illustration be helpful here? do we already have one?

@@

@@

Researchers record activities with individuals either for the purpose of a later, closer review or analysis, or simply to document and verify measurements and observations made during the collection.
Thus, the recording itself is often the core item of research, representing a single data collection, experiment, trial, visit, or case depending on the nature of the research.
However, in many cases a single session can comprise multiple videos: from serial data streams due to recording equipment producing multiple files, and/or in parallel from multiple data capture sources operating at the same time.
These primary recordings may also be augmented with other forms of timeseries data such as motion capture, eye tracking, or physiological recordings, as well as with additional point-in-time measurements, observations, or surveys, all of which have specific and critical temporal relationships with the primary data.

@@

A session object captures these relationships to represent a single data acquisition, comprising not only arbitrary files but also their absolute or relative temporal positioning.
The metaphor we use here is a timeline <good place for a screenshot> , wherein each session defines a time range, with absolute timestamps representing the exact time of data collection, or relative positioning defined by a 0 starting point of the collection (or potentially any other meaningful wall-time units) with an optional date.
Files placed within a session can then be (optionally) positioned, much as in video or audio editing software.
Although some advanced recording equipment provides automatic timestamping or synchronization than can be used, most researchers collecting multiple parallel data streams use ``sync points'' such as an electronic trigger signal, flash, or tone to align recordings.
These sync points can then be used to manually place recordings on the timeline appropriately.

@@

An object called a record serves a number of purposes around session and metadata organization. 
A record itself represents any number of variables or measurements---what would traditionally be a row of a database or flat file. Defining this component broadly supports the flexibility of the system, while giving this data a discrete place in the data schema.
For example, a record may represent an individual participant within the dataset and contain their subject identifier, gender, birth date, and any other demographic, survey, or measurement data about that individual.
A record could also represent a particular experimental task, condition, location, or outcome, and contain variables with corresponding parameters or descriptions.
Records are thus labeled and divided by the type of entity they represent.
The variables themselves are textual, numeric, date, or potentially any other scalar data as in any traditional tabular data interface, with the records of a given type then constituting the tabular rows.

@@
